---
title: "Intermediate data check"
output: 
  html_document:
    number_sections: true
editor_options: 
  chunk_output_type: console
---

```{r setup , include = FALSE, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(readxl)
library(scales)
library(readr)
library(DT)
library(ggplot2)

# clear R environment
rm(list = ls())

```

<br> 

```{r , echo = FALSE, message = FALSE, warning=FALSE}

## read files

# original data file
path <- "./01_input/01_data/01_detailed_data/detailed_data_mining.xlsx"
original <- path %>%
  excel_sheets() %>%
  set_names() %>% 
  map(~ read_excel(path = path, sheet = .x, na = c("", "-", "NA")))

# different versions of data (after conversion/harmonization etc. )
pre_con <- read_rds("./03_intermediate/01_detailed_data/all_tables_harmonized.rds")
post_con <- read_rds("./03_intermediate/01_detailed_data/all_tables_converted.rds")
conv_fac <- read_csv2("./01_input/02_lists_and_concordance_tables/source_unit_ids.csv")

# harmonized and converted data file
detailed <- read_rds("./03_intermediate/01_detailed_data/all_tables_converted.rds")

#define sheets 
sheet_min_orig <- original$minerals_ores_conce
sheet_com_orig <- original$commodities
sheet_min_pre <- pre_con$minerals_ores_conce
sheet_com_pre <- pre_con$commodities
sheet_min_post <- post_con$minerals_ores_conce
sheet_com_post <- post_con$commodities
sheet_gen <- pre_con$general
sheet_subsites <- pre_con$sub_sites



# sheets
sheet_min <- detailed$minerals_ores_conce
sheet_com <- detailed$commodities

# log file to exclude flagging of cases which have already been checked
ratio_check_log <- read_csv("./01_input/03_log_files/01_detailed_data/ratio_check_log.csv")


# create list filled up with cases where the data checks suggest a false entry
flagged_cases <- list()

```

<br> 

# Rows which are apparently double counting 

```{r, echo = FALSE, message = FALSE}

### check for double counting of data points
  ### (i.e. double entries across all relevant columns)
  ### mostly due to recording from different sources
  
  ### group twice in order to leave out those which have double entries from the same source

# sheet_min
double_min <- sheet_min %>%
  group_by(mine_fac, sub_site, type_mineral, min_ore_con, type_mining, year, mine_processing) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1) %>%
  left_join(.,
    sheet_min %>%
      select(mine_fac, sub_site, type_mineral, min_ore_con, type_mining, year, mine_processing, source_id)) %>%
  distinct() %>%
  group_by(mine_fac, sub_site, type_mineral, min_ore_con, type_mining, year, mine_processing) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1) %>%
  left_join(., sheet_min) %>%
  arrange(mine_fac, sub_site, type_mineral, min_ore_con, type_mining, year, mine_processing)

# table for Rmd output
DT::datatable(double_min, rownames = FALSE, caption = "Apparent double entries in minerals sheet")


# sheet_com
double_com <- sheet_com %>%
  group_by(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, mine_processing) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1) %>%
  left_join(.,
    sheet_com %>%
      select(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, mine_processing, source_id)) %>%
  distinct() %>%
  group_by(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, mine_processing) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1) %>%
  left_join(., sheet_com) %>%
  arrange(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, mine_processing)

# table for Rmd output
DT::datatable(double_com, rownames = FALSE, caption = "Apparent double entries in commodities sheet")





### check whether the exact same entry was inserted twice from the same source
exact_double_min <- sheet_min %>%
  group_by(mine_fac, sub_site, type_mineral, min_ore_con, type_mining, year, mine_processing, value, amount_sold) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1) %>%
  left_join(., sheet_min) %>%
  arrange(mine_fac, sub_site, type_mineral, min_ore_con, type_mining, year, mine_processing)

# table for Rmd output
DT::datatable(exact_double_min, rownames = FALSE, caption = "Exact same entry twice in minerals sheet")


exact_double_com <- sheet_com %>%
  group_by(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, mine_processing, value, amount_sold, metal_payable, grade) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1) %>%
  left_join(., sheet_com) %>%
  arrange(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, mine_processing)

# table for Rmd output
DT::datatable(exact_double_min, rownames = FALSE, caption = "Exact same entry twice in commodities sheet")




### double counting check INCLUDING double entries of the same source
### Comment: This does not necessarily mean that all entries of this output tables must be corrected, but some.

# sheet_min
double_min_incl_same_source <- sheet_min %>%
  group_by(mine_fac, sub_site, type_mineral, min_ore_con, type_mining, year, mine_processing) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1)

# table for Rmd output
DT::datatable(double_min_incl_same_source, rownames = FALSE, caption = "Double entries INCLUDING from the same source in minerals sheet (Dont necessarily have to be corrected)")


# sheet_com
double_com_incl_same_source <- sheet_com %>%
  group_by(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, mine_processing) %>%
  summarise(n = n()) %>%
  ungroup() %>%
  filter(n > 1) 

# table for Rmd output
DT::datatable(double_com_incl_same_source, rownames = FALSE, caption = "Double entries INCLUDING from the same source in commodities sheet (Dont necessarily have to be corrected)")


```

<br> 

```{r , echo = FALSE, message = FALSE}
## Add duplicates for "Ore processed" from "Ore mined" and vice versa
  ## These duplicates are only for checking and do not represent the actual duplicates introduced later after aggregation (the main reason being that without prior aggregation the duplicates might potentially create false values, especially due to different sub-sites, different mining types, and different ores) -> However, for this purpose here, it seems appropriate to include it prior aggregation as proxy values without differentiating by sub_site/type_mining/type_mineral, but only mine_fac and year.

ore_mined <- sheet_min %>%
  filter(type_mineral == "Ore mined") %>%
  distinct(mine_fac, year)

ore_proc <- sheet_min %>%
  filter(type_mineral == "Ore processed") %>%
  distinct(mine_fac, year)

# mines with reported ore mined for a year, but no ore processed for that year
mine_list <- setdiff(ore_mined, ore_proc)

sheet_min <- sheet_min %>% 
  union(
    mine_list %>%
      left_join(sheet_min %>% filter(type_mineral == "Ore mined")) %>%
      mutate(type_mineral = "Ore processed")
    )

# mines with reported ore processed, but no ore mined for that year
mine_list <- setdiff(ore_proc, ore_mined)

sheet_min <- sheet_min %>% 
  union(
    mine_list %>%
      left_join(sheet_min %>% filter(type_mineral == "Ore processed")) %>%
      mutate(type_mineral = "Ore mined")
  )

```

<br> 

```{r , echo = FALSE, message = FALSE}
## Add duplicates for "Clean coal" from "Coal mined" and vice versa
  ## Details similar to "Ore processed" and "Ore mined" above

ore_mined <- sheet_min %>%
  filter(type_mineral == "Coal mined") %>%
  distinct(mine_fac, year)

ore_proc <- sheet_min %>%
  filter(type_mineral == "Clean coal") %>%
  distinct(mine_fac, year)

# mines with reported coal mined for a year, but no Clean coal for that year
mine_list <- setdiff(ore_mined, ore_proc)

sheet_min <- sheet_min %>% 
  union(
    mine_list %>%
      left_join(sheet_min %>% filter(type_mineral == "Coal mined")) %>%
      mutate(type_mineral = "Clean coal")
  )

# mines with reported Clean coal, but no coal mined for that year
mine_list <- setdiff(ore_proc, ore_mined)

sheet_min <- sheet_min %>% 
  union(
    mine_list %>%
      left_join(sheet_min %>% filter(type_mineral == "Clean coal")) %>%
      mutate(type_mineral = "Coal mined")
  )

```

<br> 

# Mines which are in both tables, but are not joined

```{r , echo = FALSE, message = FALSE}
## check which rows from sheet_min would not be included in join of sheet_com, because variables don't fit
 ## based on six variables: mine_fac, sub_site, min_ore_con, type_mining, year, mine_processing

# matching entries
match <- intersect(
  sheet_min %>%
    filter(type_mineral %in% c("Ore processed", "Concentrate") & value != 0) %>%
    select(mine_fac, sub_site, min_ore_con, type_mining, year, mine_processing),
  sheet_com %>%
    select(mine_fac, sub_site, min_ore_con, type_mining, year, mine_processing)
)

# mine_fac which have entries in both sheets for same year, but are not matched
  # excluding those which could be matched
  # returning only the respective entries from sheet_min
non_fit <- intersect(
  setdiff(
    sheet_min %>%
      filter(type_mineral %in% c("Ore processed", "Concentrate") & value != 0) %>%
      select(mine_fac, sub_site, min_ore_con, type_mining, year, mine_processing),
    match
    ) %>%
    select(mine_fac, year),
  setdiff(
    sheet_com %>%
      select(mine_fac, sub_site, min_ore_con, type_mining, year, mine_processing),
    match
    ) %>%
    select(mine_fac, year)
  ) %>%
  left_join(.,
            setdiff(
              sheet_min %>%
                filter(type_mineral %in% c("Ore processed", "Concentrate") & value != 0) %>%
                select(mine_fac, sub_site, min_ore_con, type_mining, year, mine_processing),
              match
              )
            ) %>%
  arrange(mine_fac, desc(year))


# save no_match table as csv
write_delim(non_fit, "./04_output/01_detailed_data/03_intermediate_check/non_fit.csv", delim = ";")

# table for Rmd output
DT::datatable(non_fit)  


# comment: possible additional option would be to exclude those from sheet_min (just for the check) which are not oxide ores and at the same time have also a concentrate production reported (e.g. sulphide ore resulting in concentrate) -> because then content is likely provided for concentrate


```

<br> 

# Mines which have coordinates in sheet_general and sheet_sub_sites (not allowed, should be either/or)
```{r, echo = FALSE, message = FALSE}

### check for double entries of coordinates in sheet_general and sheet_sub_sites

general <- detailed$general
sub_sites <- detailed$sub_sites

dupli <- sub_sites %>% 
  filter(mine_fac %in% (general %>% filter(!is.na(latitude)) %>% select(mine_fac) %>% pull())) %>%
  filter(!is.na(latitude)) %>% 
  select(mine_fac) %>% 
  pull() %>% unique()

print(dupli)

```

# Mines which have only one sub-site (not allowed)
```{r, echo = FALSE, message = FALSE}
print(sheet_subsites %>% 
  group_by(mine_fac) %>% 
  summarize(n = n()) %>% 
  filter(n == 1) %>% 
  select(mine_fac) %>% 
  pull())

```

<br> 

```{r , echo = FALSE, message = FALSE}
### possible additional check: check values per mine/material for one year vs next to check for potential typos, i.e. outliers in time series


```

<br> 

```{r, echo = FALSE, message = FALSE}
### possible additional check: check for double counting between ores and concentrates


```


# Ratio checks (formerly DC3 and DC4)
These data checks are based on the relationship of the variables minerals_ores_conce.value vs commodities.value vs commodities.grade vs commodities.recovery_rate
<br>
in total, 2 ratios are calculated:
<br>

1. ratio_ore_pseudo_ore = commodities.value / (commodities.grade * commodities.recovery_rate)

2. ratio_commodity_pseudo_commodity = minerals_ores_conce.value * commodities.grade * commodities.recovery_rate

the sheets before and after the unit conversion are needed

## prepare data 
```{r , echo = FALSE, message = FALSE, results = "hide"} 

# conditions for this data check
# all 4 data points =! c(NA OR 0) AND commodities.mine_processing == NA AND type_mineral == "Ore processed"

# only include observations that fulfill conditions mentioned above
sheet_min_post_dc1_4 <- sheet_min_post %>%
  filter(!is.na(value) & value != 0 & type_mineral == "Ore processed" & is.na(mine_processing))

sheet_min_pre_dc1_4 <- sheet_min_pre %>%
  filter(!is.na(value) & value != 0 & type_mineral == "Ore processed" & is.na(mine_processing))

compcases_com <- sheet_com_post %>%
  select(value, grade, recovery_rate) %>%
  complete.cases()

sheet_com_post_dc1_4 <- sheet_com_post[compcases_com,] %>%
  filter(value != 0 & grade != 0 & recovery_rate != 0 & is.na(mine_processing))


## calculation of first ratio: ratio_ore_pseudo_ore
# combines sheet_min and sheet_com 
ratio_ore <- inner_join(sheet_com_post_dc1_4, sheet_min_post_dc1_4,  
                  by = c("mine_fac", "sub_site", "min_ore_con", "type_mining", "year"),
                  suffix = c(".com", ".min"))

# adds variables pseudo.value.min and ratio_ore_pseudo_ore
ratio_ore <- ratio_ore %>% 
              mutate(pseudo.value.min = value.com / (grade * 0.000001 * recovery_rate)) %>%
              mutate(ratio_ore_pseudo_ore = value.min / pseudo.value.min)

#include column unit which includes the original unit
ratio_ore <- ratio_ore %>% 
  left_join(sheet_min_pre_dc1_4 %>% 
              select(mine_fac, sub_site, min_ore_con, type_mining, year, unit), 
            by = c("mine_fac", "sub_site", "min_ore_con", "type_mining", "year"))
  

## calculation of second ratio: ratio_commodity_pseudo_commodity
ratio_com <- inner_join(sheet_com_post_dc1_4, sheet_min_post_dc1_4,  
                  by = c("mine_fac", "sub_site", "min_ore_con", "type_mining", "year"),
                  suffix = c(".com", ".min"))

# adds variables pseudo.value.com and ratio_commodity_pseudo_commodity
ratio_com <- ratio_com %>% 
              mutate(pseudo.value.com = value.min * grade * 0.000001 * recovery_rate) %>%
              mutate(ratio_commodity_pseudo_commodity = value.com/pseudo.value.com)

#include column unit which includes the original unit
ratio_com <- ratio_com %>% 
  left_join(sheet_com_pre %>% 
              select(mine_fac, sub_site, min_ore_con, commodity, type_mining, year, unit), 
            by = c("mine_fac", "sub_site", "min_ore_con", "commodity", "type_mining", "year"))

print("done!")
```


## Visual inspection 
### Visual inspection of the ratio ore/pseudo-ore

```{r , echo = FALSE, message = FALSE} 
datatable(ratio_ore %>% 
            select(mine_fac, sub_site, min_ore_con, type_mining, year, 
                   commodity, grade_or_yield_unit, grade, value.min, 
                   pseudo.value.min, ratio_ore_pseudo_ore),
            caption = "Full ratio_ore table")
```

Boxplot including outliers
```{r , echo = FALSE, message = FALSE, results = "hide"} 
boxplot(ratio_ore$ratio_ore_pseudo_ore)
```

Boxplot excluding outliers
```{r , echo = FALSE, message = FALSE, results = "hide"} 
boxplot(ratio_ore$ratio_ore_pseudo_ore, outline=F)
```

Histograms, untrimmed and trimmed
```{r , echo = FALSE, message = FALSE, results = "hide"} 
hist(ratio_ore$ratio_ore_pseudo_ore)

#trimming
ratio_ore_plot <- ratio_ore %>%
  filter(ratio_ore_pseudo_ore > 0.7 & ratio_ore_pseudo_ore < 1.3)

# create plots for the data checks with the unit errors 
ratio_ore_plot <- ratio_ore_plot %>%
  mutate(unit_ton = case_when(unit == "Kst" ~ "short ton", 
                              unit == "Kt" ~ "metric ton", 
                              unit == "Mt" ~ "metric ton", 
                              unit == "t" ~ "metric ton",
                              unit == "st" ~ "short ton")) 

ggplot(ratio_ore_plot, aes(x=ratio_ore_pseudo_ore)) + 
  geom_histogram(binwidth = 0.007, alpha = 0.5, color = "black")+
  theme(legend.position = c(0.23,0.8))+ 
  labs(x = "OP ratio")

ggplot(ratio_ore_plot, aes(x=ratio_ore_pseudo_ore, color = unit_ton, fill = unit_ton)) + 
  geom_histogram(binwidth = 0.007, alpha = 0.5)+
  theme(legend.position = c(0.23,0.8))+ 
  labs(x = "OP ratio")

```


```{r , echo = FALSE, message = FALSE} 
# add the extreme outliers to flagged cases
flagged_cases$dc0_1 <- ratio_ore %>% 
  filter(ratio_ore_pseudo_ore > 1.2 
         | ratio_ore_pseudo_ore < 0.8) %>%
  anti_join(ratio_check_log) %>%
  select(mine_fac, sub_site, min_ore_con, type_mining, year, 
         commodity, grade_or_yield_unit, grade, value.min, 
         pseudo.value.min, ratio_ore_pseudo_ore)

# Display for which entries the unit might be entered falsely
datatable(flagged_cases$dc0_1, caption = "Outliers with extreme ore/pseudo-ore ratios")

```

### Visual inspection of the ratio commodity/pseudo-commodity

```{r , echo = FALSE, message = FALSE} 
datatable(ratio_com %>% 
            select(mine_fac, sub_site, min_ore_con, type_mining, year, commodity, grade_or_yield_unit, grade, value.com, pseudo.value.com, ratio_commodity_pseudo_commodity),
          caption = "Full ratio_commodity table")
```

Boxplot including outliers
```{r , echo = FALSE, message = FALSE, results = "hide"}
boxplot(ratio_com$ratio_commodity_pseudo_commodity)
```

Boxplot excluding outliers
```{r , echo = FALSE, message = FALSE, results = "hide"} 
boxplot(ratio_com$ratio_commodity_pseudo_commodity, outline=F)
```

Histograms, untrimmed and trimmed
```{r , echo = FALSE, message = FALSE, results = "hide"} 
hist(ratio_com$ratio_commodity_pseudo_commodity)

# trimming
ratio_com_plot <- ratio_com %>%
  filter(ratio_commodity_pseudo_commodity > 0.7 & ratio_commodity_pseudo_commodity < 1.3)


# allocating units
ratio_com_plot <- ratio_com_plot %>%
  mutate(unit_ounce = case_when(unit == "Kg" ~ "other", 
                              unit == "t" ~ "other", 
                              unit == "Klb" ~ "other", 
                              unit == "lb" ~ "other",
                              unit == "Mlb" ~ "other",
                              unit == "Kt" ~ "other",
                              unit == "Koz" ~ "ounce",
                              unit == "oz" ~ "ounce",
                              unit == "Kozt" ~ "troy ounce",
                              unit == "ozt" ~ "troy ounce"
                              ))

ggplot(ratio_com_plot, aes(x=ratio_commodity_pseudo_commodity)) + 
  geom_histogram(binwidth = 0.007, alpha = 0.5, color = "black")+ 
  labs(x = "CP ratio")


ggplot(ratio_com_plot, aes(x=ratio_commodity_pseudo_commodity, color = unit_ounce, fill = unit_ounce)) + 
  geom_histogram(binwidth = 0.007, alpha = 0.5)+ 
  theme(legend.position = c(0.8,0.8))+ 
  labs(x = "CP ratio")

```


```{r , echo = FALSE, message = FALSE} 
#immediately add the extreme outliers to flagged cases
flagged_cases$dc0_2 <- ratio_com %>% 
  filter(ratio_commodity_pseudo_commodity > 1.2 
         | ratio_commodity_pseudo_commodity < 0.8) %>%
  anti_join(ratio_check_log) %>%
  select(mine_fac, sub_site, min_ore_con, type_mining, year, commodity, grade_or_yield_unit, grade, value.com, pseudo.value.com, ratio_commodity_pseudo_commodity)

# Display for which entries the unit might be entered falsely
datatable(flagged_cases$dc0_2, caption = "Outliers with extreme commodity/pseudo-commodity ratios")

```

## Data check: Short ton vs Metric ton (formerly DC3)
Check, if there are any cases where tonne was entered as unit, but actually short ton would be the correct unit. The other direction, i.e. entered short tonnes instead of metric tonnes, must not be checked, as metric tonnes is the default unit. Short tonnes are only entered, if the company report really specifies "Short tonnes".
All cases where the ratio ore/pseudo-ore is between 1.05 and 1.15 are flagged
The reason is that if metric tonnes are entered instead of short tonnes, the ratio is expected to be around 1.10231 (1 metric tonne = 1.10231 short tons). The actual number entered for ore_processed (which was converted in metric tonnes, but should have been converted in short tonnes) will be 10% too high 

First Visual inspection of ore/pseudo ore ratio, only including data points where the unit is some form of tonnes (i.e. tonnes, t, Kt, kt, million tonnes, short tons, long tons, etc...)
```{r , echo = FALSE, message = FALSE, results = "hide"}
# requires sheets pre and post unit conversion
# Additional Conditions: minerals_ores_conce.unit == c(tonnes, t, Kt, kt, million tonnes, long tons, etc...)

ratio_ore_dc3 <- ratio_ore %>% 
  filter(unit %in% c("t", "Kt", "Mt", "100Mt", "Gt"))

# check out ratio_ore_dc3
boxplot(ratio_ore_dc3$ratio_ore_pseudo_ore, outline=F)
ratio_ore_dc3_trimmed <- ratio_ore_dc3$ratio_ore_pseudo_ore
ratio_ore_dc3_trimmed<-ratio_ore_dc3_trimmed[ratio_ore_dc3_trimmed<1.3 & ratio_ore_dc3_trimmed>0.7]
hist(ratio_ore_dc3_trimmed, 200) 

summary(ratio_ore_dc3$ratio_ore_pseudo_ore)

# flag cases for which this confusion of short tonnes for metric tonnes could be the case
flagged_cases$dc3 <- ratio_ore_dc3 %>% 
  filter(ratio_ore_pseudo_ore > 1.05 
         & ratio_ore_pseudo_ore < 1.15) %>%
  anti_join(ratio_check_log) %>%
  select(mine_fac, sub_site, min_ore_con, type_mining, year, commodity, grade_or_yield_unit, grade, value.min, pseudo.value.min, ratio_ore_pseudo_ore)

```

Histogram with cases in and outside the acceptance region, and inside the data check region
```{r , echo = FALSE, message = FALSE} 

ratio_ore_dc3_plot <- ratio_ore_dc3 %>%
  filter(ratio_ore_pseudo_ore > 0.7 & ratio_ore_pseudo_ore < 1.3)

# normal grey histogram
ggplot(ratio_ore_dc3_plot, aes(x=ratio_ore_pseudo_ore)) + 
  geom_histogram(binwidth = 0.007, color = "black", fill = "grey")


# group ratios 
ratio_ore_dc3_plot <- ratio_ore_dc3_plot %>%
  mutate(which_check = case_when(ratio_ore_pseudo_ore <  0.8  ~ "outlier",
                                 ratio_ore_pseudo_ore >  1.2  ~ "outlier",
                                 ratio_ore_pseudo_ore >  0.95 & ratio_ore_pseudo_ore < 1.05  ~ "acceptance region",
                                 ratio_ore_pseudo_ore >  1.05 & ratio_ore_pseudo_ore < 1.15  ~ "might be entered falsely, tonnes instead of short tonnes",
                                 ratio_ore_pseudo_ore >  0.8 & ratio_ore_pseudo_ore < 0.95  ~ "no DC applied, but outside accept. reg."))


# histogram indicating acceptance and rejection region
ggplot(ratio_ore_dc3_plot, aes(x=ratio_ore_pseudo_ore, color = which_check, fill=which_check)) + 
  geom_histogram(binwidth = 0.005, alpha=0.5) + 
  theme(legend.position = c(0.23,0.8))+ 
  labs(x = "OP ratio")

# Display for which entries the unit might be entered falsely
datatable(flagged_cases$dc3, caption = "Data points which might be entered falseley, tonnes instead of short tonnes")

```



## Data check: troy ounce vs ounce (Formerly DC4)

requires sheets pre and post unit conversion
Additional Conditions: commodities.unit == c(ounce, koz, oz, 000 oz, etc.), AND commodities.grade_or_yield_unit == g/t (because if it is oz/t, it makes no sense to check if it is troy ounce instead of ounce, because if it were, oz/t would also likely refer to troy oz/t, thus making the four values fit together again, thus undetectable.)
Check: If commodities.value =~= (minerals_ores_conce.value *  commodities.grade * commodities.recovery_rate) * 0.9 --> flag
reason: if the entered.value/pseudo.value ratio is roughly 0.9, it means that the entered commodity.value is roughly 10% too small compared to the value calculated based on the other three datapoints
this is the case if ounces instead of troy ounces is entered.

First Visual inspection of commodity/pseudo-commodity ratio, only including data points where the unit is some form of ounces. 
```{r , echo = FALSE, message = FALSE, results = "hide"}

#preset the sheet_com_pre_dc4
sheet_com_pre_dc4 <- sheet_com_pre

#load the pre-conversion tibble, only where ounces or troy ounces were entered as unit
ounce_units<-conv_fac %>% #take only the units which are converted to ounces or troy ounces
  filter(unit_id %in% c("oz", "ozt", "Koz", "Kozt", "Moz", "Mozt")) %>%
  select(source_unit) %>%
  pull()

#load the pre-conversion tibble, only where not oz/t or koz/t was entered as grade
grade_units_dc4<-conv_fac %>% 
  filter(unit_cat == "grade") %>%
  filter(!(unit_id %in% c("oz/t", "ozt/t"))) %>%
  select(source_unit) %>%
  pull()

# only take observations which were converted to g/t
sheet_com_pre_dc4 <- sheet_com_pre_dc4 %>% 
  filter(unit %in% ounce_units) %>%
  filter(grade_or_yield_unit %in% grade_units_dc4) %>%
  select(mine_fac, sub_site, min_ore_con, type_mining, year, commodity, grade_or_yield_unit)


#merge the ratio_com and sheet_com_pre_dc4
ratio_com_dc4 <- inner_join(ratio_com, sheet_com_pre_dc4,  
                  by = c("mine_fac", "sub_site", "min_ore_con", "year", "type_mining", "commodity"), 
                  suffix = c("ratio_com", "sheet_com_pre"))


# check out ratio_com_dc4
boxplot(ratio_com_dc4$ratio_commodity_pseudo_commodity, outline=F)
ratio_com_dc4_trimmed <- ratio_com_dc4$ratio_commodity_pseudo_commodity
ratio_com_dc4_trimmed <- ratio_com_dc4_trimmed[ratio_com_dc4_trimmed <1.3 & ratio_com_dc4_trimmed > 0.7]
hist(ratio_com_dc4_trimmed, 200)

# fill the list flagged_cases with the cases flagged by the fourth data check
flagged_cases$dc4 <- ratio_com_dc4 %>% 
  filter(ratio_commodity_pseudo_commodity > 0.85 
         & ratio_commodity_pseudo_commodity <0.95) %>%
  anti_join(ratio_check_log) %>%
  select(mine_fac, sub_site, min_ore_con, type_mining, year, commodity, 
         grade_unit, grade, value.com, pseudo.value.com, ratio_commodity_pseudo_commodity)


```

Histogram with cases in and outside the acceptance region, and inside the data check region
```{r , echo = FALSE, message = FALSE} 

ratio_com_dc4_plot <- ratio_com_dc4 %>%
  filter(ratio_commodity_pseudo_commodity > 0.7 & ratio_commodity_pseudo_commodity < 1.3)


ratio_com_dc4_plot <- ratio_com_dc4_plot %>%
  mutate(which_check = case_when(ratio_commodity_pseudo_commodity <  0.8  ~ "outlier",
                                 ratio_commodity_pseudo_commodity >  1.2  ~ "outlier",
                                 ratio_commodity_pseudo_commodity >  0.95 & ratio_commodity_pseudo_commodity < 1.05  ~ "acceptance region",
                                 ratio_commodity_pseudo_commodity >  0.85 & ratio_commodity_pseudo_commodity < 0.95  ~ "might be entered falsely, ounces instead of troy ounces",
                                 ratio_commodity_pseudo_commodity >  0.8 & ratio_commodity_pseudo_commodity < 0.85  ~ "no check applied, but outside acceptance region",
                                 ratio_commodity_pseudo_commodity >  1.05 & ratio_commodity_pseudo_commodity < 1.2  ~ "no check applied, but outside acceptance region"))



ggplot(ratio_com_dc4_plot, aes(x=ratio_commodity_pseudo_commodity, color = which_check, fill = which_check)) + 
  geom_histogram(binwidth = 0.005, alpha=0.5)+
  theme(legend.position = c(0.8,0.8))+ 
  labs(x = "CP ratio")

# Display for which entries the unit might be entered falsely
datatable(flagged_cases$dc4, caption = "Data points which might be entered falseley, ounces instead of troy ounces")

```

Histogram commodity/pseudo-commodity ratio by grade unit before conversion
```{r , echo = FALSE, message = FALSE, results = "hide"} 

### create the plot for dc4 that also shows which unit for HG was used (building on the code from above)
sheet_com_pre_dc4_plot <- sheet_com_pre %>% 
  select(mine_fac, sub_site, min_ore_con, type_mining, year, commodity, grade_or_yield_unit)

# merge the ratio_com_plot and sheet_com_pre_dc4_plot
ratio_com_plot <- left_join(ratio_com_plot, sheet_com_pre_dc4_plot,  
                            by = c("mine_fac", "sub_site", "min_ore_con", "year", "type_mining", "commodity"), suffix = c(".ratio_com", ".sheet_com_pre"))


ggplot(ratio_com_plot, aes(x=ratio_commodity_pseudo_commodity, color = grade_or_yield_unit.sheet_com_pre, fill = grade_or_yield_unit.sheet_com_pre)) + 
  geom_histogram(binwidth = 0.007, alpha = 0.5)

```

## Other histograms: 

Histogram commodity/pseudo-commodity ratio by grade unit before conversion, without the unit "percent"
```{r , echo = FALSE, message = FALSE, results = "hide"} 
ggplot(filter(ratio_com_plot, !grade_or_yield_unit.sheet_com_pre == "pct"), aes(x=ratio_commodity_pseudo_commodity, color = grade_or_yield_unit.sheet_com_pre, fill = grade_or_yield_unit.sheet_com_pre)) + 
  geom_histogram(binwidth = 0.007, alpha = 0.5)

```

Histogram commodity/pseudo-commodity ratio by grade unit before conversion, included or not included in the data check
```{r , echo = FALSE, message = FALSE, results = "hide"} 


# make a histogram that only shows those with unit ounces or multiple
ratio_com_plot_ounces <- ratio_com_plot %>% 
  filter(!grade_or_yield_unit.sheet_com_pre == "pct") %>%
  filter(unit %in% ounce_units) %>%
  mutate(unit_grade = case_when(grade_or_yield_unit.sheet_com_pre == "g/t" ~ "grade given in g/t, included in DC4", 
                                grade_or_yield_unit.sheet_com_pre == "oz/t" ~ "grade given in oz/t, not included in DC4")) 


ggplot(ratio_com_plot_ounces, aes(x=ratio_commodity_pseudo_commodity, color = grade_or_yield_unit.sheet_com_pre, fill = grade_or_yield_unit.sheet_com_pre)) + 
  geom_histogram(binwidth = 0.007, alpha = 0.5)

ggplot(ratio_com_plot_ounces, aes(x=ratio_commodity_pseudo_commodity, color = unit_grade, fill = unit_grade)) +
  geom_histogram(binwidth = 0.007, alpha = 0.5)+
  theme(legend.position = c(0.75,0.8))+ 
  labs(x = "CP ratio")

```


Export cases which were flagged by any of the ratio data checks (i.e. 0.1, 0.2, 3 or 4)
These should, after manual checking, be either copied into the log file ratio_check_log.csv, or corrected
```{r , echo = FALSE, message = FALSE, results = "hide"} 

ratio_check_non_fit <- tibble()

for(i in 1:4){
  
  name <- flagged_cases[i] %>% names()
  
  temp <- flagged_cases[[i]] %>% 
    select(mine_fac, sub_site, year, min_ore_con, commodity) %>% 
    mutate(dc = name)
    
  
  ratio_check_non_fit <- ratio_check_non_fit %>% 
    bind_rows(temp) %>% 
    distinct()
}

ratio_check_non_fit <- ratio_check_non_fit %>% 
  arrange(dc, mine_fac, desc(year))

write_csv(ratio_check_non_fit, "./04_output/01_detailed_data/03_intermediate_check/ratio_check_non_fit.csv")

```


# Data check: check if production period matches info in sheet_general (formerly DC9)
```{r , echo = FALSE, message = FALSE} 
# General.production_start  vs minerals_ores_conce.min_ore_con 
# Conditions:
# Check: If year[ore mined] > general.production_end OR year[ore_mined] < general.production_start --> flag


# extract the mines that have entries for coal 
years_unique <- sheet_min_pre %>%
  select(mine_fac, type_mineral, year) %>%
  filter(type_mineral == "Ore mined" | type_mineral == "Coal mined") %>%
  group_by(mine_fac)

max_year <- summarise(years_unique, max = max(year)) 
min_year <- summarise(years_unique, min = min(year))


#pre-set the sheet general to only inlcude relevant variables
sheet_gen_dc9 <- sheet_gen %>% 
  select(mine_fac, production_start, production_end)


# join max_year and min_year, join with sheet_gen
production_years <- full_join(max_year, min_year, by="mine_fac")
dc9 <- full_join(sheet_gen_dc9, production_years , by="mine_fac")


dc9 <- dc9 %>% #check if there is production recorded before the entered production start, or after the production end
  mutate(prod_start_true = min - production_start) %>%
  mutate(prod_end_true = production_end - max)

# return all mines where production is recorded outside of the production start and end year
flagged_cases$dc9 <- dc9 %>% 
  filter(prod_start_true < 0 | prod_end_true < 0 ) 

# Display for which entries the unit might be entered falsely
datatable(flagged_cases$dc9, caption = "Mines for which production was recorded outside of their indicated active years")

```
# Data check: logical error in recovery rate, can only be smaller than 1 (formerly DC10)
```{r , echo = FALSE, message = FALSE} 
# commodities.recovery_rate larger than 1 
# Conditions: commodities.recovery_rate != NA AND Mine_processing == NA
# Check: if commodities.recovery_rate > 1  --> flag

flagged_cases$dc10<- sheet_com_post %>%
  filter(is.na(mine_processing))%>%
  filter(recovery_rate>1) %>%
  select(mine_fac, sub_site, min_ore_con, type_mining, year, commodity, value, recovery_rate)

# Display for which entries the unit might be entered falsely
datatable(flagged_cases$dc10, caption = "Cases where recovery rate is > 100%")

```

