---
title: "Compile Final Data"
output: html_document
editor_options: 
  chunk_output_type: console
---

###### In this script, the final data output is prepared.
###### 1. Add column facility_id to all remaining sheets
###### 2. Prepare sheet_general_sub_sites by combining the sheets general and sub_sites
###### 3. Select, rename, and arrange and round variables
###### 4. Insert join id (called minerals_id) to sheet commodities
###### 5. Export the final data in final format
###### 6. Create table "structure", storing the structure of the database and explaining variables
###### 7. Validation
###### 8. Export final data in other formats

### Read In data
```{r , echo = FALSE, message = FALSE} 

# clear R environment
rm(list = ls())

library(sf)
library(tidyverse)
library(DT)
library(RSQLite)

# read in production information
detailed <- read_rds("./03_intermediate/01_detailed_data/gaps_filled.rds")

# read in sheet_general
sheet_general <- st_read("./03_intermediate/01_detailed_data/sheet_general/07a_general.gpkg")
sheet_sub_sites <- st_read("./03_intermediate/01_detailed_data/sheet_general/sub_sites.gpkg")

# read in source_ids, material_ids, unit_ids
source_ids <- read_delim("./01_input/02_lists_and_concordance_tables/source_ids.csv", delim = ";", show_col_types = FALSE)
material_ids <- read_delim("./01_input/02_lists_and_concordance_tables/material_ids.csv", delim = ";", show_col_types = FALSE)

# read in final data structure
structure <- read_csv("./01_input/02_lists_and_concordance_tables/01_detailed_data/final_data_structure.csv", show_col_types = FALSE)

```

<br>

# 1. Add column facility_id to all remaining sheets
```{r , echo = FALSE, message = FALSE}

ids <- sheet_general %>%
  select(facility_id, mine_fac) %>% 
  st_drop_geometry() %>% 
  bind_rows(sheet_sub_sites %>% 
              select(facility_id, mine_fac, sub_site) %>% 
              st_drop_geometry())

ids_mine_fac_only <- sheet_general %>% 
  select(facility_id, mine_fac) %>% 
  st_drop_geometry()

# for the sheets processing, transport, capacity_reserves, waste, other_info and coal, it can be joined by mine_fac and sub_site
for (sheet in c("processing", "transport", "capacity_reserves", "waste", "other_info", "coal")){
  detailed[[sheet]] <- left_join(detailed[[sheet]], ids, by = c("mine_fac", "sub_site")) %>%
    relocate(facility_id, .before = mine_fac)
}

# for the sheets minerals_ores_conce, commodities, reserves, capacity it can be joined by mine_fac
for (sheet in c("minerals_ores_conce", "commodities", "reserves", "capacity")){
  detailed[[sheet]] <- left_join(detailed[[sheet]], ids_mine_fac_only, by = c("mine_fac")) %>%
    relocate(facility_id, .before = mine_fac)
}


```

<br>

# 2. Prepare sheet_general_sub_sites by combining the sheets general and sub_sites
```{r , echo = FALSE, message = FALSE}
general_combined <- list(
  sheet_general,
  sheet_sub_sites)

i = 1
for(sheet_name in c("general", "sub_sites")) {
  
  detailed[[i]] <- general_combined[[i]] %>% 
    # select cols/variables kept in final data only
    select(structure %>% 
             filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
             select(variable) %>% 
             pull()) %>% 
    # rename the cols/variables
    rename_with( ~ (structure %>%
                           filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
                           select(final_name_variable) %>% 
                           pull())) %>% 
    # relocate cols/variables in correct order 
    relocate(structure %>%
               filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
               arrange(order_cols) %>% 
               select(final_name_variable) %>% 
               pull()) %>% 
    # arrange rows in correct order
    arrange(across(all_of(structure %>%
                            filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
                            filter(!(is.na(order_rows))) %>% 
                            arrange(order_rows) %>% 
                            select(final_name_variable) %>%
                            pull()))
            )
  
  i = i + 1
}


# create sheet_general_sub_sites
detailed$general_sub_sites <- bind_rows(detailed$general, detailed$sub_sites) %>% 
  # rearrange
  arrange("facility_id") %>% 
  # relocate cols/variables in correct order 
  relocate(structure %>%
             filter(sheet == "general_sub_sites" & final_keep_variable == "yes") %>% 
             arrange(order_cols) %>% 
             select(final_name_variable) %>% 
             pull())


```

<br>

# 3. Select, rename, and arrange and round variables
```{r , echo = FALSE, message = FALSE}

#get indices of sheets in the list detailed which should be modified
sheet_numbers <- match(
  structure %>% 
    filter(final_keep_sheet == "yes" & sheet != "general_sub_sites") %>% 
    select(sheet) %>% 
    unique() %>% 
    pull(),
  detailed %>% 
    names())


#loop over the selected sheets and convert according to final data structure
for(sheet_nr in sheet_numbers){
  
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  detailed[[sheet_nr]] <- detailed[[sheet_nr]] %>% 
    # select cols/variables kept in final data only
    select(structure %>% 
             filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
             select(variable) %>% 
             pull()) %>% 
    # rename the cols/variables
    rename_with( ~ (structure %>%
                           filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
                           select(final_name_variable) %>% 
                           pull())) %>% 
    # relocate cols/variables in correct order 
    relocate(structure %>%
               filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
               arrange(order_cols) %>% 
               select(final_name_variable) %>% 
               pull()) %>% 
    # arrange rows in correct order
    # for this, first insert temporary variable consisting of the first 8 characters of facility_id (excluding sub_sites)
    arrange(across(all_of(structure %>%
                            filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
                            filter(!(is.na(order_rows))) %>% 
                            arrange(order_rows) %>% 
                            select(final_name_variable) %>% 
                            pull()))) %>%
    mutate(temp_id = substr(facility_id, 1, 8)) %>%
    arrange(temp_id) %>% 
    select(-temp_id) %>%
    # round if necessary (1. time)
    mutate_at(structure %>% 
                filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
                filter(round_cycle == 1) %>% 
                select(final_name_variable) %>% 
                pull(),
              round, 
              digits = (structure %>% 
                          filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>%
                          filter(round_cycle == 1) %>%
                          select(round_digits) %>% 
                          unique() %>% 
                          as.numeric())
              ) %>% 
    # round if necessary (2. time)
    mutate_at(structure %>% 
              filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
              filter(round_cycle == 2) %>% 
              select(final_name_variable) %>% 
              pull(),
            round, 
            digits = (structure %>% 
                        filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>%
                        filter(round_cycle == 2) %>%
                        select(round_digits) %>% 
                        unique() %>% 
                        as.numeric())
            ) %>% 
    # replace unnecessary newlines
    mutate(across(where(is.character), str_replace_all, "[\r\n]" , ""))

  # create unique id corresponding to the row-number 
  # not for table general, as there the facility_id is the primary key
  if(curr_sheet_name != "general_sub_sites") {
    detailed[[sheet_nr]] <- detailed[[sheet_nr]] %>% 
      add_column(id = 1:nrow(detailed[[sheet_nr]]), .before = "facility_id")
  }

  print(c("Sheet", curr_sheet_name, "done"))
}


```

<br>

# 4. Insert join id (called id_minerals) to sheet commodities
```{r , echo = FALSE, message = FALSE}

commodities_joined <- left_join(
  detailed$commodities,
  detailed$minerals_ores_conce %>% 
    filter((type %in% c("Ore processed", "Concentrate"))), # exclude where type_mining is "Ore mined" or "NM mineral"
  by = c("facility_id", "material", "year", "mine_processing", "reporting_period"), 
  suffix = c("_commodities", "_minerals")
)


detailed$commodities <- detailed$commodities %>% 
  left_join(commodities_joined %>% 
              select(id_commodities, id_minerals),
            by = c("id" = "id_commodities"))

```

<br>

# 5. Create tables source_ids, and material_ids
```{r , echo = FALSE, message = FALSE}
# source_ids
source_ids <- source_ids %>% 
  rename(sources = source, 
         source_urls = source_url)


# material_ids

# table with material_ids
material_ids_singular <- material_ids %>% 
  filter(!(is.na(material_id))) %>%
  select(material_id, material_name, material_category, material_category_2, material_id_agg, material_name_agg)


# table with the aggregated material ids, which will be added to the material_ids table
material_ids_agg <- material_ids %>% 
  filter(!(is.na(material_id_agg))) %>% 
  group_by(material_id_agg) %>% 
  summarize(
    material_id_agg = unique(material_id_agg),
    material_name_agg = unique(material_name_agg),
    material_category = "aggregate",
    material_category_2 = paste(unique(material_category_2), "aggregate"),
    materials_contained = paste(unique(material_id), collapse = "; ") %>% 
      str_replace("; NA|NA ;|NA | NA|NA", "")
    ) %>% 
  rename(material_id = material_id_agg,
         material_name = material_name_agg) 


material_ids <- material_ids_singular %>% 
  bind_rows(material_ids_agg)

```

<br>

# 6. Create table structure which stores the structure of the database and explains variables
```{r , echo = FALSE, message = FALSE}

all_variables <- structure %>% 
  select(final_name_sheet, final_name_variable, description, supposed_type, example_entry, references_table) %>%
  filter(!(is.na(description))) %>%
  rename(table = final_name_sheet,
         variable = final_name_variable,
         example = example_entry,
         type = supposed_type) %>% 
  arrange(factor(table, levels = c("general", "coal", "minerals", "commodities", "processing",
                                   "transport", "reserves", "capacity", "waste", "ownership", "material_ids", "source_ids")))


```

<br>

# 7. Validation

# 7.1 Validation: check for NA Values where they are allowed
```{r , echo = FALSE, message = FALSE}
na_tables <- list()

#loop over the selected sheets and convert according to final data structure
i = 1
for(sheet_nr in sheet_numbers){
  
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  
  complete_cases <- detailed[[sheet_nr]] %>% 
    # select cols/variables where no na is allowed
    select(structure %>%
             filter(sheet == curr_sheet_name & na_allowed == "no") %>% 
             select(final_name_variable) %>% 
             pull()) %>% 
    # check if an NA value exists in the remaining dataframe
    complete.cases()
  
  if(length(unique(complete_cases)) == 1 & unique(complete_cases) == TRUE) {
    print(paste0("Sheet ", curr_sheet_name, " is complete"))
  } else {
    print(paste0("Warning: Sheet ", curr_sheet_name, " is NOT complete and contains NA values where none are allowed!"))
     na_tables[[i]] <- detailed[[sheet_nr]][complete_cases == FALSE,] %>% 
           select(structure %>%
             filter(sheet == curr_sheet_name & na_allowed == "no") %>% 
             select(final_name_variable) %>% 
             pull()) %>% 
       left_join(detailed$general_sub_sites %>% select(facility_id, facility_name)) %>% 
       relocate(facility_name, .after = facility_id)
     i = i+1
  }
}

# you have to manually check how many items the na_tables list has 
# uncomment them to display data table on Rmd output
# length(na_tables)
# datatable(na_tables[[1]])
# datatable(na_tables[[2]])
# datatable(na_tables[[3]])
# datatable(na_tables[[4]])
# datatable(na_tables[[5]])
# datatable(na_tables[[6]])

```

<br>

# 7.2 Validation: Display graphical summaries to check for outliers/incorrect data
```{r , echo = FALSE, message = FALSE}
# 1. plot histogram value_tonnes for several tables
# ggplot(data = detailed$minerals_ores_conce, aes(x = value_tonnes)) + 
#   geom_histogram(bins = 100)
# 
# detailed$minerals_ores_conce %>% 
#   left_join(detailed$general_sub_sites %>% select(facility_id, facility_name)) %>% 
#   relocate(facility_name, .after = facility_id) %>% 
#   arrange(value_tonnes %>% desc) %>% 
#   head()

```

<br>

# 7.3 Validation: validate with SQL schema provided by PGModeler
```{r , echo = FALSE, message = FALSE}

# still to do

```

<br>

# 8. Export the final data in final format
```{r , echo = FALSE, message = FALSE}

# create the directory for storing final data
dir.create("./04_output/01_detailed_data/08_final_data/final_format_txt/", showWarnings = FALSE, recursive = TRUE)
dir.create("./04_output/01_detailed_data/08_final_data/final_format_csv/", showWarnings = FALSE, recursive = TRUE)

# table general is separate, as it is exported as a geopackage (TXT)
st_write(detailed$general_sub_sites %>% st_as_sf(), "./04_output/01_detailed_data/08_final_data/final_format_txt/general.gpkg", 
          append = FALSE)

# table general is separate, as it is exported as a geopackage (CSV)
st_write(detailed$general_sub_sites %>% st_as_sf(), "./04_output/01_detailed_data/08_final_data/final_format_csv/general.gpkg", 
          append = FALSE)

#loop over rest of the sheets in detailed and export as TXT and CSV
for(sheet_nr in sheet_numbers[-length(sheet_numbers)]){
  
  #define current sheet name
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  # print(curr_sheet_name)
  # print(detailed[[sheet_nr]])
  
  #create output path
  final_sheet_name <- structure %>% 
    filter(sheet == curr_sheet_name) %>% 
    select(final_name_sheet) %>% 
    unique() %>% 
    pull()
  
  path_txt <- paste0("./04_output/01_detailed_data/08_final_data/final_format_txt/", final_sheet_name, ".txt")
  path_csv <- paste0("./04_output/01_detailed_data/08_final_data/final_format_csv/", final_sheet_name, ".csv")

  #export txt
  write_delim(detailed[[sheet_nr]], 
              file = path_txt, 
              na = "", 
              delim = "\t")
  
  #export csv
  write_csv(detailed[[sheet_nr]], 
              file = path_csv, 
              na = "")
  

}

# export source_ids, material_ids as TXT files
write_delim(source_ids, "./04_output/01_detailed_data/08_final_data/final_format_txt/source_ids.txt", na = "", delim = "\t")
write_delim(material_ids, "./04_output/01_detailed_data/08_final_data/final_format_txt/material_ids.txt", na = "", delim = "\t")

# export source_ids, material_ids as CSV files
write_csv(source_ids, "./04_output/01_detailed_data/08_final_data/final_format_csv/source_ids.csv", na = "")
write_csv(material_ids, "./04_output/01_detailed_data/08_final_data/final_format_csv/material_ids.csv", na = "")

# export variable descriptions
write_csv(all_variables, "./04_output/01_detailed_data/08_final_data/variables_descriptions.csv", na = "")

```

<br>

# 9. Export final data in other formats
```{r , echo = FALSE, message = FALSE}

# create the directory for storing other formats
dir.create("./04_output/01_detailed_data/08_final_data/other_formats/", showWarnings = FALSE, recursive = TRUE)

write_rds(detailed, "./04_output/01_detailed_data/08_final_data/other_formats/detailed_final.rds")

```

<br>


# 9.1 Export sheet_general, sheet_sub_sites separately
```{r , echo = FALSE, message = FALSE} 

# export sheet_general_sub_sites as TXT
write_delim(detailed$general_sub_sites %>%
            select(-geom),
          "./04_output/01_detailed_data/08_final_data/other_formats/general.txt", na = "", delim = "\t")

# export sheet_general_sub_sites as CSV
write_csv(detailed$general_sub_sites %>%
            select(-geom),
          "./04_output/01_detailed_data/08_final_data/other_formats/general.csv", na = "")

# test table general
# general_test <- st_read("./04_output/01_detailed_data/08_final_data/final_format/general.gpkg")
# 
# general_test$geom
# 
# general_test

```


# 9.2 Export all tables to single geopackage file 
```{r , echo = FALSE, message = FALSE}

# write table general to geopackage file (database)
st_write(detailed$general_sub_sites %>% st_as_sf(),
         dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data.gpkg", 
         layer = "general", append = FALSE)


#loop over rest of the sheets in detailed and export to database
for(sheet_nr in sheet_numbers[-length(sheet_numbers)]){
  
  #define current sheet name
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  # print(curr_sheet_name)
  # print(detailed[[sheet_nr]])
  
  #create output path
  final_sheet_name <- structure %>% 
    filter(sheet == curr_sheet_name) %>% 
    select(final_name_sheet) %>% 
    unique() %>% 
    pull()
  
  # add to database
  st_write(detailed[[sheet_nr]],
         dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data.gpkg", 
         layer = final_sheet_name, append = FALSE)
  
}


# add source_ids, material_ids to database
st_write(source_ids,
       dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data.gpkg", 
       layer = "source_ids", append = FALSE)

st_write(material_ids,
       dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data.gpkg", 
       layer = "material_ids", append = FALSE)

```


# 9.3 Export all tables to single geopackage file, excluding facilities without coordinates
```{r , echo = FALSE, message = FALSE}
# data preparation: exclude facilities without coordinates, and filter all sheets accordingly
fac_w_coord_list <- detailed$general_sub_sites %>% 
  filter(geom %>% st_is_empty() == FALSE) %>% 
  st_drop_geometry() %>% 
  select(facility_id) %>% 
  pull()


# write table general to geopackage file (database)
st_write(detailed$general_sub_sites %>% 
           st_as_sf() %>% 
           filter(facility_id %in% fac_w_coord_list),
         dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data_georef_only.gpkg", 
         layer = "general", append = FALSE)


#loop over rest of the sheets in detailed and export to database
for(sheet_nr in sheet_numbers[-length(sheet_numbers)]){
  
  #define current sheet name
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  
  #create output path
  final_sheet_name <- structure %>% 
    filter(sheet == curr_sheet_name) %>% 
    select(final_name_sheet) %>% 
    unique() %>% 
    pull()
  
  # create final table by filtering out facilities without coordinates
  temp <- detailed[[sheet_nr]] %>% 
    filter(facility_id %in% fac_w_coord_list)
  
  # add to database
  st_write(temp,
         dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data_georef_only.gpkg", 
         layer = final_sheet_name, append = FALSE)
  
}


# add source_ids, material_ids to database
st_write(source_ids,
       dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data_georef_only.gpkg", 
       layer = "source_ids", append = FALSE)

st_write(material_ids,
       dsn = "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data_georef_only.gpkg", 
       layer = "material_ids", append = FALSE)

# checking / testing
# 
# st_layers("./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data.gpkg")
# st_layers("./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data_georef_only.gpkg")
# 
# test_old <- st_read(
#   "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data.gpkg",
#   layer = "general")
# 
# test <- st_read(
#   "./04_output/01_detailed_data/08_final_data/other_formats/mine_level_data_georef_only.gpkg",
#   layer = "minerals")
# 
# test %>% st_geometry_type() %>% unique()
# test %>% st_is_empty() %>% unique()

```

