---
title: "Compile Final Data"
output: html_document
editor_options: 
  chunk_output_type: console
---

###### In this script, the final data output is prepared.
###### 1. Add column mine_id to all remaining sheets
###### 2. Prepare sheet_general_sub_sites by combining the sheets general and sub_sites
###### 3. Select, rename, and arrange and round variables
###### 4. Insert join id (called minerals_id) to sheet commodities
###### 5. Export the final data in final format
###### 6. Validation
###### 7. Export final data in other formats

### Read In data
```{r , echo = FALSE, message = FALSE} 

# clear R environment
rm(list = ls())

library(sf)
library(tidyverse)
library(DT)

# read in production information
detailed <- read_rds("./03_intermediate/01_detailed_data/gaps_filled.rds")

# read in sheet_general
sheet_general <- st_read("./03_intermediate/01_detailed_data/sheet_general/general.gpkg")
sheet_sub_sites <- st_read("./03_intermediate/01_detailed_data/sheet_general/sub_sites.gpkg")

# read in source_ids, material_ids, unit_ids
source_ids <- read_delim("./01_input/02_lists_and_concordance_tables/source_ids.csv", delim = ";")
material_ids <- read_delim("./01_input/02_lists_and_concordance_tables/material_ids.csv", delim = ";")
unit_ids <- read_delim("./01_input/02_lists_and_concordance_tables/unit_ids.csv", delim = ";")

# read in final data structure
structure <- read_csv("./01_input/02_lists_and_concordance_tables/01_detailed_data/final_data_structure.csv")

```

<br>

# 1. Add column mine_id to all remaining sheets
```{r , echo = FALSE, message = FALSE}

ids <- sheet_general %>%
  select(mine_id, mine_fac) %>% 
  st_drop_geometry() %>% 
  bind_rows(sheet_sub_sites %>% 
              select(mine_id, mine_fac, sub_site) %>% 
              st_drop_geometry())

ids_mine_fac_only <- sheet_general %>% 
  select(mine_id, mine_fac) %>% 
  st_drop_geometry()


# for the sheets processing, transport, capacity_reserves, waste, other_info and coal, it can be joined by mine_fac and sub_site
for (i in c(5,6,8,9,11)){
  detailed[[i]] <- left_join(detailed[[i]], ids, by = c("mine_fac", "sub_site")) %>%
    relocate(mine_id, .before = mine_fac)
}

# for the sheets minerals_ores_conce, commodities it can be joined by mine_fac
for (i in c(3,4,12,13)){
  detailed[[i]] <- left_join(detailed[[i]], ids_mine_fac_only, by = c("mine_fac")) %>%
    relocate(mine_id, .before = mine_fac)
}
```

<br>

# 2. Prepare sheet_general_sub_sites by combining the sheets general and sub_sites
```{r , echo = FALSE, message = FALSE}
general_combined <- list(
  sheet_general %>% tibble(),
  sheet_sub_sites %>% tibble())

i = 1
for(sheet_name in c("general", "sub_sites")) {
  
  detailed[[i]] <- general_combined[[i]] %>% 
    # select cols/variables kept in final data only
    select(structure %>% 
             filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
             select(variable) %>% 
             pull()) %>% 
    # rename the cols/variables
    rename_with( ~ (structure %>%
                           filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
                           select(final_name_variable) %>% 
                           pull())) %>% 
    # relocate cols/variables in correct order 
    relocate(structure %>%
               filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
               arrange(order_cols) %>% 
               select(final_name_variable) %>% 
               pull()) %>% 
    # arrange rows in correct order
    arrange(across(all_of(structure %>%
                            filter(sheet == sheet_name & final_keep_variable == "yes") %>% 
                            filter(!(is.na(order_rows))) %>% 
                            arrange(order_rows) %>% 
                            select(final_name_variable) %>%
                            pull()))
            )
  
  i = i + 1
}


# create sheet_general_sub_sites
detailed$general_sub_sites <- bind_rows(detailed$general, detailed$sub_sites) %>% 
  arrange("facility_id")


```

<br>

# 3. Select, rename, and arrange and round variables
```{r , echo = FALSE, message = FALSE}

#get indices of sheets in the list detailed which should be modified
sheet_numbers <- match(
  structure %>% 
    filter(final_keep_sheet == "yes") %>% 
    select(sheet) %>% 
    unique() %>% 
    pull(),
  detailed %>% 
    names())

#loop over the selected sheets and convert according to final data structure
for(sheet_nr in sheet_numbers){
  
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  detailed[[sheet_nr]] <- detailed[[sheet_nr]] %>% 
    # select cols/variables kept in final data only
    select(structure %>% 
             filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
             select(variable) %>% 
             pull()) %>% 
    # rename the cols/variables
    rename_with( ~ (structure %>%
                           filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
                           select(final_name_variable) %>% 
                           pull())) %>% 
    # relocate cols/variables in correct order 
    relocate(structure %>%
               filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
               arrange(order_cols) %>% 
               select(final_name_variable) %>% 
               pull()) %>% 
    # arrange rows in correct order
    arrange(across(all_of(structure %>%
                            filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
                            filter(!(is.na(order_rows))) %>% 
                            arrange(order_rows) %>% 
                            select(final_name_variable) %>% 
                            pull()))) %>% 
    # round if necessary (1. time)
    mutate_at(structure %>% 
                filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
                filter(round_cycle == 1) %>% 
                select(final_name_variable) %>% 
                pull(),
              round, 
              digits = (structure %>% 
                          filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>%
                          filter(round_cycle == 1) %>%
                          select(round_digits) %>% 
                          unique() %>% 
                          as.numeric())
              ) %>% 
    # round if necessary (2. time)
    mutate_at(structure %>% 
              filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>% 
              filter(round_cycle == 2) %>% 
              select(final_name_variable) %>% 
              pull(),
            round, 
            digits = (structure %>% 
                        filter(sheet == curr_sheet_name & final_keep_variable == "yes") %>%
                        filter(round_cycle == 2) %>%
                        select(round_digits) %>% 
                        unique() %>% 
                        as.numeric())
            )

    # create unique id corresponding to the row-number 
  # not for table general, as there the facility_id is the primary key
  if(curr_sheet_name != "general_sub_sites") {
    detailed[[sheet_nr]] <- detailed[[sheet_nr]] %>% 
      add_column(id = 1:nrow(detailed[[sheet_nr]]), .before = "facility_id")
  }

  print(c("Sheet", curr_sheet_name, "done"))
}


```

<br>

# 4. Insert join id (called id_minerals) to sheet commodities
```{r , echo = FALSE, message = FALSE}

commodities_joined <- left_join(
  detailed$commodities,
  detailed$minerals_ores_conce %>% 
    filter((type %in% c("Ore processed", "Concentrate"))), # exclude where type_mining is "Ore mined" or "NM mineral"
  by = c("facility_id", "material", "year", "mine_processing", "reporting_period"), 
  suffix = c("_commodities", "_minerals")
)


detailed$commodities <- detailed$commodities %>% 
  left_join(commodities_joined %>% 
              select(id_commodities, id_minerals),
            by = c("id" = "id_commodities"))


```

<br>

# 5. Create tables source_ids, material_ids, unit_ids 
```{r , echo = FALSE, message = FALSE}
# source_ids does not change

unit_ids <- unit_ids %>%
  select(-wu_former) %>% 
  rename(unit_category = unit_cat)

material_ids <- material_ids %>% 
  select(material_id, material_name, material_category, material_category_2) %>% 
  rename(material = material_id)

```

<br>

# 6. Validation
```{r , echo = FALSE, message = FALSE}

# 1. check if there are NA values where none can be

na_tables <- list()

#loop over the selected sheets and convert according to final data structure
i = 1
for(sheet_nr in sheet_numbers){
  
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  
  complete_cases <- detailed[[sheet_nr]] %>% 
    # select cols/variables where no na is allowed
    select(structure %>%
             filter(sheet == curr_sheet_name & na_allowed == "no") %>% 
             select(final_name_variable) %>% 
             pull()) %>% 
    # check if an NA value exists in the remaining dataframe
    complete.cases()
  
  if(length(unique(complete_cases)) == 1 & unique(complete_cases) == TRUE) {
    print(paste0("Sheet ", curr_sheet_name, " is complete"))
  } else {
    print(paste0("Warning: Sheet ", curr_sheet_name, " is NOT complete and contains NA values where none are allowed!"))
     na_tables[[i]] <- detailed[[sheet_nr]][complete_cases == FALSE,] %>% 
           select(structure %>%
             filter(sheet == curr_sheet_name & na_allowed == "no") %>% 
             select(final_name_variable) %>% 
             pull()) %>% 
       left_join(detailed$general_sub_sites %>% select(facility_id, facility_name)) %>% 
       relocate(facility_name, .after = facility_id)
     i = i+1
  }
}

# you have to manually check how many items the na_tables list has 
# uncomment them to display data table on Rmd output
# length(na_tables)
# datatable(na_tables[[1]])
# datatable(na_tables[[2]])
# datatable(na_tables[[3]])
# datatable(na_tables[[4]])
# datatable(na_tables[[5]])
# datatable(na_tables[[6]])


```

<br>

# 7. Export the final data in final format
```{r , echo = FALSE, message = FALSE}

# create the directory for storing final data
dir.create("./04_output/01_detailed_data/08_final_data/final_format/", showWarnings = FALSE, recursive = TRUE)

# table general is separate, as it is exported as a geopackage
st_write(detailed$general_sub_sites, "./04_output/01_detailed_data/08_final_data/final_format/general.gpkg", 
          append = FALSE)

#loop over rest of the sheets in detailed and export as CSVs
for(sheet_nr in sheet_numbers[-length(sheet_numbers)]){
  
  #define current sheet name
  curr_sheet_name <- detailed[sheet_nr] %>% names()
  print(curr_sheet_name)
  print(detailed[[sheet_nr]])
  
  #create data table in markdown
  datatable(head(detailed[[sheet_nr]], rownames = FALSE, caption = curr_sheet_name))

  #create output path
  final_sheet_name <- structure %>% 
    filter(sheet == curr_sheet_name) %>% 
    select(final_name_sheet) %>% 
    unique() %>% 
    pull()
  
  path <- paste0("./04_output/01_detailed_data/08_final_data/final_format/", final_sheet_name, ".csv")
  
  #export
  write_csv(detailed[[sheet_nr]], path, na = "")

}

# export source_ids, material_ids, unit_ids
write_csv(source_ids, "./04_output/01_detailed_data/08_final_data/final_format/source_ids.csv", na = "")
write_csv(material_ids, "./04_output/01_detailed_data/08_final_data/final_format/material_ids.csv", na = "")
write_csv(unit_ids, "./04_output/01_detailed_data/08_final_data/final_format/unit_ids.csv", na = "")


```

<br>

# 8. Export final data in other formats
```{r , echo = FALSE, message = FALSE}
# 
# # create the directory for storing other formats
# dir.create("./04_output/01_detailed_data/08_final_data/other_formats/", showWarnings = FALSE, recursive = TRUE)

```

<br>

# 9. Export sheet_general, sheet_sub_sites separately
```{r , echo = FALSE, message = FALSE} 
# 
# # datatable for Rmd output
# datatable(head(detailed$general %>% st_drop_geometry()), rownames = FALSE, caption = "sheet_general")
# datatable(head(detailed$sub_sites %>% st_drop_geometry()), rownames = FALSE, caption = "sheet_sub_sites")
# datatable(head(detailed$general_sub_sites %>% st_drop_geometry()), rownames = FALSE, caption = "sheet_general_sub_sites")
# 
# if(export_general == TRUE) {
#   
#   # create the directory for storing sheet_general
#   dir.create("./04_output/01_detailed_data/08_final_data/other_formats/general/", showWarnings = FALSE, recursive = TRUE)
#   
#   
#   # export sheet_general as geopackage
#   st_write(detailed$general, "./04_output/01_detailed_data/08_final_data/other_formats/general/sheet_general.gpkg", append = FALSE)
#   
#   # export sheet_sub_sites as geopackage
#   st_write(detailed$sub_sites, "./04_output/01_detailed_data/08_final_data/other_formats/general/sheet_sub_sites.gpkg", append = FALSE)
#   
#   
#   # export sheet_general as CSV
#   write_csv(detailed$general %>% 
#               st_drop_geometry() %>% 
#               select(-GADM_level, -n_features), 
#             "./04_output/01_detailed_data/08_final_data/other_formats/general/sheet_general.csv", na = "")
#   
#   # export sheet_sub_sites as CSV
#   write_csv(detailed$sub_sites %>% 
#               st_drop_geometry(),
#             "./04_output/01_detailed_data/08_final_data/other_formats/general/sheet_sub_sites.csv", na = "")
#   
#   # export sheet_general_sub_sites as CSV
#   write_csv(detailed$general_sub_sites %>% 
#               st_drop_geometry() %>% 
#               select(-GADM_level, -n_features), 
#             "./04_output/01_detailed_data/08_final_data/other_formats/general/sheet_general_sub_sites.csv", na = "")
# }
# 

```

<br>

# 10. Table Mineral + Coal
```{r , echo = FALSE, message = FALSE}
# 
# #table with minerals and coal 
# sheet_min_coal <- bind_rows(detailed$minerals_ores_conce, detailed$coal)
# datatable(head(sheet_min_coal), rownames = FALSE, caption = "sheet_min_coal")
# write_csv(sheet_min_coal, "./04_output/01_detailed_data/08_final_data/other_formats/minerals_coal.csv", na = "")

```
